---
title: "Hyperspectral Games"
author: "..."
date: "27/8/2020" 
output:
  html_document:
 # ioslides_presentation: 
    highlight: tango
    incremental: yes
    smaller: yes
    transition: faster
    widescreen: yes
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<!-- ## Shiny Presentation -->

<!-- The following single character keyboard shortcuts enable alternate display modes: -->

<!-- 'f': enable fullscreen mode  -->
<!-- 'w': toggle widescreen mode  -->
<!-- 'o': enable overview mode  -->
<!-- 'h': enable code highlight mode  -->
<!-- 'p': show presenter notes -->

<!-- Pressing Esc exits all of these modes -->

## Notes

Code of this file (index.Rmd) available in GitHub [here](https://github.com/fpirotti/HyperspectralGames){target=_blank}

The white target is a bit in the way, next time better to put it on top so it can be cropped out easily.  

The cloth used as background is not "opaque" in the sense of Lambertian reflector (perfect diffusion)... it is translucid, reflecting the light source with directionality.

Image was cropped to save space and memory.

The two sensors do not provide perfect overlap with a simple rescaling, because sensor geometry is different (focal length, lens distorsions etc...) you can check by 

## Load libraries

It is a good habit in coding in R to use the namespace, i.e. include, before the function, the library it comes from, separated by "::"  e.g. the "raster" library is used to read the data, so write "raster::brick(....)".

```{r message=FALSE,  echo=TRUE }

library(shinyWidgets)
library(raster)
library(rgdal) 
library(ggplot2) 
library(plotly)
library(readr) 
```

## Read the data

We read the two images that were previously cut (see file index.R which has some preprocessing steps). Hint - try to use rgdal package which is much faster than the raster package for many things.

```{r echo=TRUE }

myImages<-list()

swir.file<-"/archivio/home/pirotti/Google Drive/RAD_3451-1_SWIR_384me_SN3155_22000us_2020-02-11T144756_raw_rad.tif"
vnir.file<-"/archivio/home/pirotti/Google Drive/RAD_3451-1_VNIR_1800_SN00852_22000us_2020-02-11T144756_raw_rad.tif"
  
myImages[['SWIR']]<-raster::brick(swir.file)
myImages[['VNIR']]<-raster::brick(vnir.file)


extent(myImages[['VNIR']]) <- extent(myImages[['SWIR']])

### here we extract wavelengths from the band names
swir.bands.wavelengths<-readr::parse_number(names(myImages[['SWIR']])) 
vnir.bands.wavelengths<-readr::parse_number(names(myImages[['VNIR']])) 

### here we put them in a string with a suffix with the sensor name, we will use this later to populate sliders for choosing the band combinations to plot and RGB image.
bands<-c( paste0("VNIR-", names(myImages[['VNIR']])),
          paste0("SWIR-", names(myImages[['SWIR']]) ) )

## trick to order the bands all together
all.wavelengths<-c(vnir.bands.wavelengths, swir.bands.wavelengths )
##  order function gives us a list of integers of the element position in the array in alphabetical order.
ord<-order(all.wavelengths)

## here we order the bands  
bands.ordered<-bands[ord]
#bands.ordered.w<-all.wavelengths[ord]

## an ordere list of sensor-band wavelength to add to the 3 sliders
bands.ordered.list<-as.list(bands.ordered)
  
``` 

## Plot Channels

Assign 3 bands to give to screen channels... Red - Green - Blue from top to bottom slider respectively - wait around 20 seconds for resampling and plot redrawing - takes more time if you choose the VNIR as it is a bigger image.

The code is not visible, check index.Rmd  in GitHub [here](https://github.com/fpirotti/HyperspectralGames){target=_blank}

```{r plot1}

    shinyWidgets::sliderTextInput("bandR", 
                                  label=NULL,
                                   width="100%",
                                  choices = bands.ordered.list,
                                  selected = bands.ordered.list[[350]] )
    shinyWidgets::sliderTextInput("bandG", 
                                  label=NULL,
                                   width="100%",
                                  choices = bands.ordered.list,
                                  selected = bands.ordered.list[[320]])
    shinyWidgets::sliderTextInput("bandB", 
                                  label=NULL,
                                   width="100%",
                                  choices = bands.ordered.list,
                                  selected = bands.ordered.list[[280]])

 plotOutput("plot1", click = "plot_click")


 output$plot1 <- renderPlot({
   
  req(input$bandR, input$bandG, input$bandB)
   
  p<-list()
  curType<-NULL
  for(i in c("bandR", "bandG", "bandB")){
    img_and_band <- stringr::str_split(input[[i]], "-")[[1]]
 
    p[[i]]<-myImages[[ img_and_band[[1]] ]][[   img_and_band[[2]] ]]
     
    if( !is.null(curType) && curType!=img_and_band[[1]] ){
      p[[i]]<-raster::resample( p[[i]], p$bandR )
    }
    curType<-img_and_band[[1]]
  }
  
  final.raster<-raster::stack(p)

  plotRGB(final.raster,
      r = 1, g = 2, b = 3, 
      scale=255,
      stretch = "hist")

})

```

&nbsp;
 

```{r plot2}

 plotlyOutput("graph1")
output$graph1 <- renderPlotly({
  
    swir.values<-gdalUtils::gdallocationinfo(swir.file, input$plot_click$x,
                                input$plot_click$y, valonly=T)
    vnir.values<-gdalUtils::gdallocationinfo(vnir.file, input$plot_click$x,
                                input$plot_click$y, valonly=T)
    
    vnir.values.num <- as.numeric(vnir.values)
    swir.values.num <- as.numeric(swir.values)
    
    x.click<-as.integer(input$plot_click$x)
    y.click<-as.integer(input$plot_click$y)
    if(!shiny::isTruthy(x.click)) x.click<-0
    if(!shiny::isTruthy(y.click)) y.click<-"0 - click image above."
    p <- plot_ly(type="scatter",mode="markers")
    p <- layout(p,  xaxis=list(title = "Wavelength (nm)"),
                    yaxis=list(title = "Radiance at sensor ( ~reflectance )"), 
                    title= sprintf("Spectral signature at X=%s Y=%s",  
                                 x.click, y.click ) )
    p <- add_trace(p, y=vnir.values.num, x=vnir.bands.wavelengths, name="VNIR",mode="markers")
    p <- add_trace(p, y=swir.values.num, x=swir.bands.wavelengths, name="SWIR",mode="markers")
    p
    
  }) 


```

## OK let's classify!

We will use only the SWIR because it is a lighter image and takes less time to process in real time and we do not want to be bored waiting for image to classify.

Grab some training pixels by choosing the class and start clicking.


... to be continued ....





```{r  }
  plotOutput("plot2", click = "plot_click2")

selectInput("myClasss", label = "Choose class and click to add trainers", choices = c("Background", "White target", "Healty Plant", "Stressed Plant"))
 output$plot2 <- renderPlot({
   
  plotRGB(myImages$SWIR,
      r = 100, g = 60, b = 30, 
      scale=255,
      stretch = "hist")
 })
 
  observeEvent({input$plot_click2$x
                input$plot_click2$y}, {
    
  })
 

```

